---
title: "Mobility Map"
author: "Sara"
date: "March 20, 2017"
output: 
  html_document:
    toc: true
    toc_depth: 3
---
```{r setup, include=FALSE}
# knitr options
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(compare)
library(sf)
library(leaflet)
library(leaflet.extras)
library(readr)

#files
data_file <- "~/Downloads/mobility_data.csv"
cz_center_file <- "~/Downloads/cz_centroids.rds"
geom_file <- "~/Downloads/geom_data.rds"

# Albers projection for 48 contiguous US states
US_ALBERS <-
  "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0"
  # GPS long-lat projection
WGS84 <- 
  "+proj=longlat +datum=WGS84 +no_defs"

  # Colors
COLORS <- c(
  "#FFFFE1",
  "#FFFFC4",
  "#FFFAAC",
  "#FFD858",
  "#FFD94F",
  "#FEB93A",
  "#FF7B00",
  "#D64700",
  "#BA1B00",
  "#990200"
)

NA_COLOR <- "#f2f2f2"

#codes for alaska and hawaii
CZS_AK_HI <- c(
  "34101",
  "34102",
  "34103",
  "34104",
  "34105",
  "34106",
  "34107",
  "34108",
  "34109",
  "34110",
  "34111",
  "34112",
  "34113",
  "34114",
  "34115",
  "34701",
  "34702",
  "34703",
  "35600"
)

DESCRIPTION <- 
  c(stringr::str_c('<p style="font-weight: normal">',
                   'This is <b style="font-weight: bold">very low</b>',
                   ' compared with</br>the rest of the country.</p>'),
    stringr::str_c('<p style="font-weight: normal">',
                   'This is <b style="font-weight: bold">below average</b>',
                   '</br>compared with the rest of the</br>country.</p>'),
    stringr::str_c('<p style="font-weight: normal">',
                   'This is <b style="font-weight: bold">about average</b>',
                   '</br>compared with the rest of the</br>country.</p>'),    
    stringr::str_c('<p style="font-weight: normal">',
                   'This is <b style="font-weight: bold">above average</b>',
                   '</br>compared with the rest of the</br>country.</p>'),
    stringr::str_c('<p style="font-weight: normal">',
                   'This is <b style="font-weight: bold">very high</b>',
                   ' compared</br>with the rest of the country.</p>'))

CZ_MOB_DOMAIN <- c(34, 42.11, 45.07, 48, 60.8)
```

Store data in a tibble:

```{r, message=FALSE}
#read in col names
names <- read_csv(file = data_file, skip = 15, n_max = 0) 
#read in data
data <- read_csv(file = data_file, skip = 17, 
                 col_names = attributes(names)$names) %>% 
  slice(2:n()) %>% 
  dplyr::select(cz = `Commuting Zone`,
         cz_name = `CZ Name`,
         intercept = `Income Age 26 Intercept`,
         slope = `Income Age 26 Slope`)
```

Calculate absolute upward mobility measurement, as defined in Chetty et al. (2014):

```{r}
data <-
  data %>% 
  mutate(mobility = intercept + .25*slope)
```

Compute mean absolute mobility per commuting zone:

```{r}
data <-
  data %>% 
  group_by(cz, cz_name) %>% 
  summarise(mean_mobility = mean(mobility*100, na.rm = TRUE)) %>% 
  ungroup()
```

Read in geometry data:
```{r}
geom_data <- read_rds(geom_file) 

geom <-
  geom_data$q3 %>% 
  st_transform(crs = WGS84) %>% 
  mutate(cz = as.integer(cz_1990)) %>% 
  dplyr::select(cz, geometry)
```

Join mobility data to geometry data:
```{r}
geom2 <-
  geom %>% 
  left_join(data, by = "cz") %>% 
  st_transform(crs = WGS84)
```

Get state lines data:
```{r}
state_lines <- 
  geom_data$q5 %>% 
  st_transform(crs = WGS84)
```

Separate state name from commuting zone name:
```{r}
data <-
  data %>% 
  separate(cz_name, into = c("cz_name", "state"), sep = ", ")
```

The following function formats the name and mobility measure for a hover-over label.
```{r}
hover_format <- function(czname, mobility) {
    if (is.na(mobility)) {
    t <- '<span style="font-weight: normal">No data.</span>'
  } else {
    t <- 
      stringr::str_c('<p style="font-weight: normal">',
                     'In the ', 
                     '<b style="font-weight: bold">', czname, '</b>', 
                     ' area, absolute upward', 
                     '</br>',
                     ' mobility is ', 
                     '<b style="font-weight: bold">', 
                     htmltools::htmlEscape(format(mobility, digit = 3)),
                     '.</p>',
                     DESCRIPTION[findInterval(mobility, CZ_MOB_DOMAIN) + 1])
  }
  htmltools::HTML(t)
}
```

Read in the centroid data (geometry for the center a given commuting zone):
```{r}
cz_center <- read_rds(cz_center_file) 
```

Color scale for the map:
```{r}
color_scale <- function(x) {
    scale_fill_gradientn(colours = COLORS, 
                       values = seq(max(data$mean_mobility, na.rm = TRUE), 
                                     min(data$mean_mobility, na.rm = TRUE)) %>% 
                                scales::rescale(),
                       na.value = NA_COLOR,
                       breaks = 26:52)$palette(scales::rescale(x)) %>% 
    stringr::str_replace_na(replacement = "#F3F3F3")
}
```

Labeling function for each commuting zone.
```{r}
label_format <- function(czname) {
  mob <- data %>% 
    filter(cz_name == czname) %>% 
    .$mean_mobility
    t <- 
      stringr::str_c('<p>',
                     '<b style="font-weight: bold">', czname, '</b>', 
                     '</br>',
                     '<b style="font-weight: normal">', 
                     htmltools::htmlEscape(format(mob, digit = 3)), '</b>',
                     '</p>')
  htmltools::HTML(t)
}
```

Plot:
```{r}
leaf_crs <- leafletCRS(crsClass = "L.Proj.CRS", 
                  code = "EPSG:42303",
                  proj4def = US_ALBERS, 
                  resolutions = c(5500, 5500 / 5.5)) 

leaf_options <- leafletOptions(crs = leaf_crs, maxZoom = 2)

leaflet(options = leaf_options,
        width = "1000px",
        height = "650px") %>% 
  addPolygons(data = geom2, 
              group = "polygons",
              fillOpacity = 1,
              fillColor = ~color_scale(mean_mobility),
              color = "white", 
              weight = 0.1, 
              opacity = 1,
              highlightOptions = highlightOptions(color = "black",
                                                  weight = 1,
                                                  opacity = 1,
                                                  bringToFront = TRUE,
                                                  sendToBack = TRUE),
              label = ~map2(cz_name, mean_mobility, hover_format),
              labelOptions = 
                labelOptions(direction = "down", 
                             opacity = 1, 
                             textOnly = TRUE,
                             offset = c(0, 25),
                             style = list('border' = '1px solid #BBBBBB',
                                         'border-radius' = '0px',
                                         'font-size' = '10px',
                                         'background-color' = '#FFFFFF',
                                         'line-height' = '110%'))) %>% 
     addLabelOnlyMarkers(data = cz_center,
                         group = "labels",
                         label = ~map(czname, label_format),
                         labelOptions =
                               labelOptions(direction = "bottom",
                                            noHide = 'T',
                                            textOnly = T,
                                            textsize = '10px',
                                            style = list(
                                              'text-align' = 'center',
                                              'line-height' = '105%',
                                              'font-size' = '9pt'
                                            )
                                            )) %>%
  addPolylines(data = state_lines, 
               group = "state_lines",
               weight = 1,
               color = "gray") %>% 
  groupOptions(group = "labels", zoomLevels = 1) %>% 
  addLegend(color = color_scale(seq(26, 52, 3)),
              labels = as.character(seq(26, 52, 3)),
              opacity = 1,
              title = "Absolute upward mobility",
              position = "bottomright") %>% 
  setMapWidgetStyle(list(background = "white")) %>% 
  setMaxBounds(-131, 20, -68, 48.42) 
```

